---
title: "Homework2"
author: "Isabel Nelson"
date: "9/24/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```

### Problem 1: Read + clean data

Read in Mr. Trashwheel dataset:   
```{r import_trash}
trashwheel_df <- read_xlsx(
  "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = "Mr. Trash Wheel",
  range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% #this only gets rid of NA values in the dumpster col
  mutate(
    sports_balls = round(sports_balls), 
    sports_balls = as.integer(sports_balls) #can modify the previous variable we just mutated
  )
```

Read in precipitation data for 2017 and 2018:   
```{r import_prec}
precip2018_df <- read_xlsx(
  "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = "2018 Precipitation",
  skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip2017_df <- read_xlsx(
  "./Data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = "2017 Precipitation",
  skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Combine two precipitation datasets: 
```{r combine}
month_df <-
  tibble(
    month = 1:12, 
    month_name = month.name
  )

precip_df <-
  bind_rows(precip2017_df, precip2018_df)

left_join(precip_df, month_df, by = "month")

#Extract out variable for the answers to problem 1 to use in paragraph below. 
```

This dataset has information from the Mr. Trashwheel trash collector in Baltimore, Maryland. When trash enters the harbor, the trashwheel collects it and puts it into a dumpster. The dataset has information on year, month, and trash collected, including some specific types of trash. There are a total of `r nrow(trashwheel_df)` rows in the final dataset. Additional data sheets include month precipitation data. 




